Implications of afforestation policies for global markets

### 2024-02-27 paper.Rmd
Update paper.Rmd

When examining the likely impacts of the baseline changes alone, it is estimated that by 2033, significant deforestation would have occured in many regions of the world. For example, in Brazil...

Then, we examined what it would take given the pressures of a still growing global population and an increasing appetite for foodstuffs for the practice of deforestation to end, on net, globally. Results suggest that it would require a tax of roughly 70.3 percent to be assessed on the value of converted land for the deforestation rate to fall globally to roughly zero. The economic impact of such a policy measure would not be limited to the cost of the tax, it would also change impact food production and global trade flows. For example, with this policy in place, most regions of the world would see significant declines in the amount of agricultural land in 2033 relative to the baseline scenario where land conversion was not taxed. For the United States, there would be 2.1 percent less agricultural land (Table 1). Larger decreases would happen in other major agricultural production regions such as China (-2.5 percent), South Asia (-8.0 percent), Brazil (-8.5 percent), and Russia (-9.3 percent).

Table 1 qo-Land
				∆ (%)
East Asia			-14.6
Central America			-14.5
North Africa & the Middle East	-10.3
Russia				-9.3
Southeast Asia			-9.0
Brazil				-8.5
Rest of South America		-11.7
South Asia			-8.0
Canada				-7.9
West & Central Europe 		-6.2
Oceania				-5.9
Sub-Saharan Africa		-4.1
China				-2.5
United States			-2.2
Japan				17.5

These reductions in the amount of agricultural land would then impact the quantity of agricultural goods that could be produced. For example, the 14.6 percent decrease in agrciultural land in East Asia and the 9.0 percent decrease in agrciultural land in Southeast Asia would result in large declines in percentage terms in agricultural production of 5.1 percent and 2.0 percent, respectively (Table 2). Surprisingly, Brazil and the United States, despite their dereases in available agricultural land compared to the baseline 2033 scenario, are estimated to see increased agricultural output of 0.2 percent and 0.6 percent, respectively. This may be tied to the sharp declines in other regions prompting greater intensity of production to satisfy external demand.

Table 2 qo-Ag
				∆ (%)
				Agriculture
East Asia			-5.1
Southeast Asia			-2.0
Central America			-2.0
South Asia			-1.1
North Africa & the Middle East	-0.9
Sub-Saharan Africa		-0.3
China				-0.1
Russia				0.1
Brazil				0.2
Rest of South America		-1.2
West & Central Europe 		0.4
United States			0.6
Oceania				0.7
Canada				1.3
Japan				1.4

When we examine how this impacts exports of agricultural goods, we find that exports from Japan and the United States grow the most, increasing 8.8 percent and 3.5 percent, respectively (table 3). The regions where agricultural exports drop most significantly are East Asia, South Asia, Southeast Asia, and Central America who see decreases of 14.5 percent, 5.9 percent, 5.1 percent,and 5.0 percent, respectively.

Table 3 qxw-Ag
				∆ (%)
				Agriculture
Japan				8.78
United States			3.49
Canada				3.13
China				1.96
Oceania				1.67
Russia				1.52
West & Central Europe 		1.16
Brazil				1.03
Rest of South America		-2.67
Sub-Saharan Africa		0.02
North Africa & the Middle East	-1.52
Central America			-5.02
Southeast Asia			-5.14
South Asia			-5.94
East Asia			-14.45

>>>>

When examining the likely impacts of the baseline changes alone, it is estimated that between 2023 and 2033, 144.2 million hectares (3.6%) of global forests will be converted to agricultural lands with the largest decline expected to happen in Sub-Saharan Africa, which is estimated to lose 39.7 million hectares (7.0%) of its forestland (Table 1). Southeast Asia, Brazil, and the rest of South America also see forestland shrink significantly, with decreases of 22.6 million hectares (9.3%), 19.0 million hectares (3.8%), and 23.5 million hectares (6.7%), respectively. Conversely, the United States is estimated to add 1.3 million hectares (0.5%) to its forestland.

Table 1. Forestland by region in 2023 and estimated change by 2033
	Russia	Sub-Saharan Africa Brazil Canada Rest of South America	United States Southeast Asia Oceania EU	China Central America South Asia Other Europe North Africa & the Middle East Japan East Asia Total 
Forestland in 2023	838.1	564.7	498.2	412.8	348.5	248.4	243.9	199.2	186.7	164.6	94.6	47.8	40.2	33.8	23.7	14.5	3,959.5
Forestland in 2033	828.7	525.0	479.2	407.9	325.0	249.7	221.3	196.1	175.5	164.8	91.2	42.6	36.1	30.9	26.9	14.6	3,815.3
∆ (ha)			-9.5	-39.7	-19.0	-5.0	-23.5	1.3	-22.6	-3.1	-11.1	0.1	-3.4	-5.3	-4.1	-2.9	3.2	0.1	-144.2
∆ (%)			-1.1	-7.0	-3.8	-1.2	-6.7	0.5	-9.3	-1.5	-6.0	0.1	-3.6	-11.0	-10.1	-8.5	13.7	0.7	-3.6

Against this backdrop, we examined what it would take to end, on net, the practice of deforestation globally. Results suggest that it would require a tax of roughly 70.3 percent to be assessed on the value of converted land for the deforestation rate to fall globally to zero. The economic impact of such a policy measure would not be limited to the cost of the tax, it would also impact food production and global trade flows. For example, with this policy in place, most regions of the world would see significant declines in the amount of agricultural land in 2033 relative to the baseline scenario where land conversion was not taxed. For the United States, there would be 2.1 percent less agricultural land (Table 2). Larger decreases would happen in other major agricultural production regions such as China (-2.5 percent), South Asia (-8.0 percent), Brazil (-8.5 percent), and Russia (-9.3 percent).

Table 2. Percentage change in the amount of agricultural land in 2033 with deforestation mitigation policy in place (Source: GTAP qo-Land)
				∆ (%)
East Asia			-14.6
Central America			-14.5
North Africa & the Middle East	-10.3
Russia				-9.3
Southeast Asia			-9.0
Brazil				-8.5
Rest of South America		-11.7
South Asia			-8.0
Canada				-7.9
West & Central Europe 		-6.2
Oceania				-5.9
Sub-Saharan Africa		-4.1
China				-2.5
United States			-2.2
Japan				17.5

These reductions in the amount of agricultural land would then impact the quantity of agricultural goods that could be produced. For example, the 14.6 percent decrease in agrciultural land in East Asia and the 9.0 percent decrease in agrciultural land in Southeast Asia would result in large declines in percentage terms in agricultural production of 5.1 percent and 2.0 percent, respectively (Table 3). Surprisingly, Brazil and the United States, despite their dereases in available agricultural land compared to the baseline 2033 scenario, are estimated to see increased agricultural output of 0.2 percent and 0.6 percent, respectively. This may be tied to the sharp declines in other regions prompting greater intensity of production to satisfy external demand.

Table 3. Percentage change in the amount of agricultural land in 2033 with deforestation mitigation policy in place (Source: GTAP qo-Land)qo-Ag
				∆ (%)
				Agriculture
East Asia			-5.1
Southeast Asia			-2.0
Central America			-2.0
South Asia			-1.1
North Africa & the Middle East	-0.9
Sub-Saharan Africa		-0.3
China				-0.1
Russia				0.1
Brazil				0.2
Rest of South America		-1.2
West & Central Europe 		0.4
United States			0.6
Oceania				0.7
Canada				1.3
Japan				1.4

When we examine how this impacts exports of agricultural goods, we find that exports from Japan and the United States grow the most, increasing 8.8 percent and 3.5 percent, respectively (Table 4). The regions where agricultural exports drop most significantly are East Asia, South Asia, Southeast Asia, and Central America who see decreases of 14.5 percent, 5.9 percent, 5.1 percent,and 5.0 percent, respectively.

Table 4. Percentage change in the amount of agricultural land in 2033 with deforestation mitigation policy in place (Source: GTAP qo-Land)qxw-Ag
				∆ (%)
				Agriculture
Japan				8.78
United States			3.49
Canada				3.13
China				1.96
Oceania				1.67
Russia				1.52
West & Central Europe 		1.16
Brazil				1.03
Rest of South America		-2.67
Sub-Saharan Africa		0.02
North Africa & the Middle East	-1.52
Central America			-5.02
Southeast Asia			-5.14
South Asia			-5.94
East Asia			-14.45
